A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans

### 2023-07-13 manuscript/manuscript.tex
ms updates

Our work extends theory from \textcite{Santiago2016-mu}, which suggests both
strong and weak negative selection can be be accommodated with a quantitative
genetic model of linked selection. These linked selection models stem from
\textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
fitness variation generates weak genetic draft. At the individual level
Robertson considered, selection generates autocorrelation as offspring from
large families tend to beget many descendents themselves (and likewise with
small families) when fitness is heritable. This same autocorrelation occurs at
the genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as
the perturbations to a neutral allele's trajectory from its particular fitness
background tend to occur in the same direction across generations until the
background recombines off. These models quantify the total impact of the
autocorrelation generated by weak draft in terms of what we think of as a
\emph{draft-effective} population size $N_d$. The key insight is that in the
long-run, steady, weak draft created by the presence of additive genetic
fitness variance $V_A > 0$ acts like an extra variance in offspring number
would under pure drift \parencite{Wright1938-tv}. However, due to
autocorrelation, the cumulative effect of this fitness variance is inflated by
by a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the
expected total variance in reproductive success a neutral mutation experiences
over its lifetime in a system with weak selection draft.

We define the draft-effective population size $N_d$ by including the total
additional variance created by heritable fitness in Wright's
(\citeyear{Wright1938-tv}) equation,

@@

Equation \eqref{eq:main_Ne} is general, since different modes of selection and
linkage can be accommodated by different expressions for the inflation factor
$Q^2$ \parencite{Santiago1995-hx,Santiago1998-bs}. When fitness variation has a
multiplicative polygenic basis, as is often assumed for genome-wide negative
selection processes, the draft asymptotic effective population size experienced
by an arbitrary neutral site under the influence of all linked regions is,

\begin{align}
    \label{eq:polygenic_Ne}
    N_d \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
\end{align}

where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Materials Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. Under equilibrium levels of fitness variation, the
autocorrelation function for a neutral allele associated with segment $i$ is
$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
the segment and $\kappa_i$ is the rate that the associated fitness variance
decays due to selective dynamics. Then, the cumulative autocorrelation is,

@@

Under any selection model, the genic fitness variance created by a new mutation
(at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx \nicefrac{s^2}{N}$.
For the entire population of $2N$ chromosomes, the fresh variance from mutation
each generation in segment $i$ is $V_{m,i} \approx U_is^2$ where $U_i = 2\mu
L_i$ is the diploid mutation rate per generation within the segment. Under
mutation-selection balance assumed by BGS, an $L_i$-basepair segment has genic
variance $V_{a,i}^{BGS} \approx U_i s$ (see Supplementary Materials Equation
\ref{eq:va_bgs}) and thus $\kappa_i^{BGS} = s$. Substituting $V_{a,i}^{BGS}$
and $\kappa_i^{BGS}$ in Equation \eqref{eq:simple_Ne} and simplifying, we have

@@

Given that modeling the interplay of mutation, drift, and draft under both weak
and strong selection has proven to be difficult problem, we first sought to
verify the SC16 theory and our genome-wide implementation of it with a three
levels of simulations.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_1.pdf} \caption{Theory
        compared to simulation results. (A) The predicted reduction factor
        under classic B theory (dark gray line) and the diploid SC16 model
        (colored lines corresponding to mutation rate) compared to average
        reduction across XXX simulation replicates (points). (B) The predicted
        ratchet rate under the SC16 model scaled by mutation rate (colored
        lines) compared to the ratchet estimated from simulation (points). When
        $2Ns>1$, the ratchet rate is near zero. (C) The genic variance from
        simulations (points) against the predicted variance under the SC16
        model (colored lines). As the ratchet begins to click, the genic
        variance is decreased from the level expected under strong BGS (dashed
        line). (D, bottom) The mean squared error (MSE) between
        whole-chromosome simulations and predicted classic B (dots), new B'
        (solid), and locally-rescaled B' (dashed) for different mutation rates
        (colors). The dashed horizontal line is the approximate theoretic
        minimum MSE. (D, top) The build up of negative linkage disequilibria
        around $2Ns=1$ in whole-chromosome simulations in bottom panel. (E) The
        average B map from 100 chromosome 10 simulation replicates (gray)
    against different predictions, for parameters that correspond to $2Ns < 1$,
$2Ns = 1$, and $2Ns > 1$. The chromosome shows the density of conserved sites
and recombination map used in simulations. }
  \label{fig:figure-1}
\end{figure}

@@

\subsection*{Annotation Model Comparison}

@@

Overall, our $R_\text{LOCO}^2$ estimates suggest our negative selection model
explains up to 67\% of out-sample variance in diversity of the megabase scale,
even though our method assumes constant demography and homogeneous mutation
rates along the genome. A worthwhile question is: how much variation
\emph{could} we expect to fit at this scale? Given that selection alters
genealogies in ways beyond just decreasing mean pairwise coalescence time and
populations have non-constant demography, an exact analytic answer is
intractable. However, we can get an approximate idea if we assume that the
residual variance $\sum_b (\pi_b - \widehat{\pi}_b)^2$ is determined entirely
by the expected neutral coalescence noise around the expected coalescence time
$2B(b)N$. This can be be found analytically, plugging-in our predictions for
$B(b)$. This allows us to calculate $R_\text{coal}^2$ to ballpark the theoretic
variance that is capable of being explained, assuming this coalescence-only
noise process (see Method XXX). We note that selection is expected to
\emph{decrease} the variance in coalescence times beyond a rescaled effective
population size implies; our $R_\text{coal}^2$ would be an underestimate under
selection.

@@

Early genome-wide classic BGS models fit well, but lead to unusually high
estimates of the mutation rate \parencite{McVicker2009-ax}. This lead to the
hypothesis that these models could be absorbing the signal of positive
selection \parencite{Enard2014-kz}, though other work has found a limited role
for hitchhiking \parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022}. Still,
a central benchmark of fitting genome-wide models is that mutation rate
estimates are in accordance with pedigree-based estimates (XXX).

We find across all populations, our mutation rate estimates from CADD-based
models are roughly concordant with pedigree-based estimates (Figure
\ref{fig:figure-4}A), consistent with recent work \parencite{Murphy2022-sj}.
Our full-track CADD 6\% model estimates the mutation rate as $\widehat{\mu} =
1.46 \times 10^{-8}$ for YRI, $1.53 \times 10^{-8}$ for CEU, and $1.51 \times
10^{-8}$ for CHB reference samples (Supplementary Table \ref{supp:tbl-r2}). As
expected, the sparse-track CADD model mutation rate estimates are identical
between the B' and classic BGS methods (Figure \ref{fig:figure-4}A top row).

However, mutation rate estimates for feature-based annotation models show
several issues. First, mutation rate estimates under our implementation of
classic BGS theory are an order of magnitude below the expected range (Figure
\ref{fig:figure-4}A top row). We observe similar behavior when we use the
classic BGS model to fit full-coverage annotation models (Figure
\ref{fig:figure-4}A bottom row). This behavior is consistent with classic BGS
theory being unable to fit the DFE to features under weak constraint (e.g.
introns, UTRs, and the ``other" feature), and consequently severally
underestimating the mutation rate.

Second, we noticed that across all populations and sparse/full tracks, the CADD
6\% model consistently led to slightly higher mutation rates than the CADD 8\%
model (Figure \ref{fig:figure-4}A bottom row; Supplementary Table
\ref{supp:tbl-r2}). This pattern was observed in Murphy et al. too (2023;
Appendix 1, Figure 16). This behavior is consistent with a non-identifiability
issue between a slightly higher per-basepair mutation rate and annotation
tracks that contain more conserved sequence. This is expected from theory,
since both classic BGS and SC16 models only depend on mutation rate through the
compound parameter $\mu L$, where $L$ is the length of the conserved segment.
Even though our method is much more robust to including non-conserved regions
like introns, we still observe this non-identifiability issue.

Finally, we note that mutation rate estimates in the Feature Priority model are
unreasonably high ($\widehat{\mu} \approx 4.2 \times 10^{-8}$), reminiscent of
the high mutation rate estimates found under McVicker et al.'s
(\citeyear{McVicker2009-ax}) model. While both our and Murphy et al.'s CADD and
PhastCons-based models alleviate this issue, it is worth discussing why this
could occur. We gain some insight from comparing the estimated mutation rates
between our Feature and PhastCons Priority annotation models, which both
contain the exact same number of basepairs, but these are assigned to different
feature classes based on the priority of overlapping features. That one of
these models is our best-fitting model, and the other our worst indicates the
sensitivity these models have when features classes have heterogeneous DFEs.
CADD-based models fit better in part due to their fine-scale resolution of
selective effects across the genome. While ideally we would fit a CADD model
with different features corresponding to the different percentiles of
pathogenicity, these features are on the basepair scale and thus too
memory-intensive for our method to currently accommodate.

@@

    \caption{(A) Observed and predicted diversity under the CADD 6\% full-track
        model. Once scaled by average diversity, populations differ little in
        their predicted diversity levels (colors). Additionally, we show
        summaries of CADD density and recombination rate along the chromosome
        below. (B) Predicted $B$ and observed $\pi$ for each window. The
        residual variance is in the vicinity of the levels expected from the
        $R_\text{LOCO}^2$ calculation. (C) Evidence that negative selection
    signal remains in CADD 6\% residuals (YRI shown). Residuals are plotted
against the average LoF selection coefficient across genes in a megabase window
(estimated from \cite{Agarwal2023-un}).}

@@

While we have previously shown that out-sample predicted diversity matches
observed diversity (Section XXX), we also investigated whether there were
regions of the genome that were particularly poor-fitting. Comparing predicted
against observed diversity along chromosomes (see Supplementary Materials
Section \ref{supp:chrom-fit}) we find a close correspondence consistent with
the high $R_\text{LOCO}^2$. However, we find a few large (tens of megabases)
regions with systematically poorer fit (Supplementary Materials Figure
\ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one such
region on the short arm of chromosome 2, from 30 Mbp to 60Mbp. Interestingly,
predicted diversity closely follows the peaks and troughs of this region. We
note that a small region within this stretch had been found by a genome-wide
scan for associative overdominance \parencite{Gilbert2020-aw}.

Another model diagnostic is to inspect whether observations are systematically
different from predictions. We confirm the finding of \textcite{Murphy2022-sj}
that regions predicted to experience little reduction in diversity due to
background selection have higher diversity than expected (Figure
\ref{fig:figure-5}B, orange line). Their work suggested that this could reflect
ancient introgression between archaic humans and ancestors of contemporary
humans. Finally, the variance around observed and predicted diversity levels
falls very close to what we would expect under the theoretic
coalescence-noise-only expectation ($R_\text{coal}^2$).

Earlier, we found that the goodness-of-fit and DFE estimates of our
feature-based annotation models are sensitive to how basepairs are grouped into
feature classes. Our models are in essence fitting a single distribution across
sites within a feature class, yet there may be local heterogeneity among sites
that is poorly captured by this approach. We investigated this possibility by
looking for lingering selection signal in our model residuals. First, we
inspected whether there was a relationship between the fraction of CADD 2\% and
6\% basepairs and the residual across megabase windows (Supplementary Figure
\label{suppfig:resid-cadd}), finding a negative significant relationship in
both cases. Moreover, our model over-predicted diversity in windows containing
more CADD 2\% basepairs moreso than CADD 6\%, consistent with heterogeneity in
site pathogenicity being poorly fit by our model. However, the total residual
variance explained is $R^2 = 0.4\%$ and $R^2 = 0.8\%$ for the CADD 2\% and 6\%
tracks respectively, suggesting only modest amount of selection signal remains.

@@

Since our B' method is based upon quantitative genetic models of linked
selection that explicitly incorporates the substitution process, fitting our
model to genomic data leads to predictions: (1) the substitution rates per
segment, (2) the fitness variance per segment, and (3) the total number of
strongly deleterious mutations per diploid genome per generation. Given
predicted substitution rates for each feature can be compared to estimated
sequence divergence between species, this offers a qualitative check our model
is producing sensible results.

To compare our models' predicted substitution rates against observed
divergence, we estimated sequence divergence on the human lineage using a
multiple alignment of five primates for each feature in our feature-based
models (Methods XXX). Since we have observed that mutation rate estimates from
our method can be unreliable (and simulation results confirm this), we
predicted substitution rates under a fixed mutation rate of $\mu = 1.5 \times
10^{-8}$, rather than the maximum likelihood estimates. Unfortunately, a
careful comparison between our predictions and observed divergence rates is
hindered by considerable uncertainty in the rate of the molecular clock,
generation times, and the human-chimpanzee divergence time. We assume a
generation time of 29 years, and calculate the sequence divergence implied by
our predicted substitution rates over a range of divergence times, from 6 Mya
to 12 Mya \parencite{Moorjani2016-tb,Nachman2000-te,Yi2002-pw,Steiper2006-xx}.

Our predictions are qualitatively consistent with the observed divergence along
the human lineage (Figure \ref{fig:figure-6}). As expected, the predicted
substitution rate in features unlikely to be under selective constraint is very
close to the mutation rate. Throughout, we report our substitution rates as a
percent relative to the total mutation rate, $\mu$ (here fixed to $1.5 \times
10^{-8}$). In our PhastCons Priority model, introns and UTRs are estimated to
have have substitution rates that are $93.6\%$ of the mutation rate, and the
``other" feature is estimated at $97.8\%$.  When we use the constrained grid,
these change to $97.85\%$ and $99.98\%$ respectively. In our PhyloFit
estimates, the divergence rate in introns is $92.44\%$, 5' UTRs $75.5\%$, and
3' UTRs $72.65\%$. By contrast, our predicted substitution rates for PhastCons
regions are much lower: $0.48\%$ for the default grid, and $1.65\%$ for the
constrained grid. For comparison, the PhyloFit estimates for PhastCons regions
is much higher, at $51.23\%$. Since this would include beneficial substitutions
that would not be predicted by our model, this suggests a very high proportion
of beneficial substitutions XXX check

>>>>

Our work extends theory from \textcite{Santiago2016-mu}, which suggests both
strong and weak negative selection can be be accommodated with a quantitative
genetic model of linked selection. These linked selection models stem from
\textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
fitness variation generates weak genetic draft. At the individual level
Robertson considered, selection generates autocorrelation as offspring from
large families tend to beget many descendents themselves (and likewise with
small families) when fitness is heritable. This same autocorrelation occurs at
the genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as
the perturbations to a neutral allele's trajectory from its particular fitness
background tend to occur in the same direction across generations until the
background recombines off. These models quantify the total impact of the
autocorrelation generated by weak draft in terms of what we think of as a
\emph{draft-effective} population size $N_d$. The key insight is that in the
long-run, steady, weak draft created by the presence of additive genetic
fitness variance $V_A > 0$ acts like an extra variance in offspring number
would under pure drift \parencite{Wright1938-tv}. However, due to
autocorrelation, the cumulative effect of this fitness variance is inflated 
by a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the
expected total variance in reproductive success a neutral mutation experiences
over its lifetime in a system with weak selection draft.

We can define the draft-effective population size $N_d$ by including the total
additional variance created by heritable fitness in Wright's
(\citeyear{Wright1938-tv}) equation for effective population size,

@@

Equation \eqref{eq:main_Ne} is general, since different modes of selection and
linkage can be accommodated by different expressions for $V_A$ and the
inflation factor $Q^2$ \parencite{Santiago1995-hx,Santiago1998-bs}. When
fitness variation has a multiplicative polygenic basis, as is often assumed for
genome-wide negative selection processes, the draft asymptotic effective
population size experienced by an arbitrary neutral site under the influence of
all linked regions is,

\begin{align}
    \label{eq:polygenic_Ne}
    N_d \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
\end{align}
%
where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Materials Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. Under equilibrium levels of fitness variation, the
autocorrelation function for a neutral allele associated with segment $i$ is
$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
the segment and $\kappa_i$ is the rate that the associated fitness variance
decays due to selective dynamics. Then, the cumulative autocorrelation is,

@@

Under any selection model, the additive genic fitness variance created by a new
mutation (at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx
\nicefrac{s^2}{N}$. For the entire population of $2N$ chromosomes, the fresh
variance from mutation each generation in segment $i$ is $V_{m,i} \approx
U_is^2$ where $U_i = 2\mu L_i$ is the diploid mutation rate per generation
within the segment. Under mutation-selection balance assumed by BGS, an
$L_i$-basepair segment has genic variance $V_{a,i}^{BGS} \approx U_i s$ (see
Supplementary Materials Equation \ref{supp-eq:va_bgs}) and thus $\kappa_i^{BGS}
= s$. Substituting $V_{a,i}^{BGS}$ and $\kappa_i^{BGS}$ in Equation
\eqref{eq:polygenic_Ne} and simplifying, we have

@@

Given that modeling the interplay of mutation, drift, and draft under both weak
and strong selection has proven to be difficult problem, we first sought to
verify the SC16 theory and our genome-wide extension of this theory with three
levels of simulations: forward simulations of negative selection in a region,
chromosome-scale forward simulations of negative selection, and simulations of
a ``synthetic genome" to test our composite-likelihood method based on this
theory.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_1.pdf} \caption{Theory
        compared to simulation results. (A) The predicted reduction factor
        under classic B theory (dark gray line) and the diploid SC16 model
        (colored lines corresponding to mutation rate) compared to average
        reduction across 10,000 simulation replicates (points). (B) The predicted
        ratchet rate under the SC16 model scaled by mutation rate (colored
        lines) compared to the ratchet estimated from simulation (points). When
        $2Ns>1$, the ratchet rate is near zero. (C) The genic variance from
        simulations (points) against the predicted variance under the SC16
        model (colored lines). As the ratchet begins to click, the genic
        variance is decreased from the level expected under strong BGS (dashed
        line). (D, bottom) The mean squared error (MSE) between
        whole-chromosome simulations and predicted classic B (dots), new B'
        (solid), and locally-rescaled B' (dashed) for different mutation rates
        (colors). The dashed horizontal line is the approximate theoretic
        minimum MSE. (D, top) The build up of negative linkage disequilibria
        around $2Ns=1$ in whole-chromosome simulations in bottom panel. (E) The
        average B map from 100 chromosome 10 simulation replicates (gray)
    against different predictions, for parameters that correspond to $2Ns < 1$,
$2Ns = 1$, and $2Ns > 1$. The chromosome shows the density of conserved sites
and recombination map used in simulations. }
  \label{fig:figure-1}
\end{figure}

@@

\subsection*{Annotation Model Comparison}
\label{sec:annotation}

@@

Overall, our $R_\text{LOCO}^2$ estimates suggest our negative selection model
explains up to 67\% of out-sample variance in diversity of the megabase scale,
even though our method assumes constant demography and homogeneous mutation
rates along the genome. A worthwhile question is: how much variation
\emph{could} we expect to fit at this scale? Given that selection alters
genealogies in ways beyond just decreasing mean pairwise coalescence time and
populations have non-constant demography, an exact analytic answer is
intractable. However, we can get an approximate idea if we assume that the
residual variance $\sum_b (\pi_b - \widehat{\pi}_b)^2$ is determined entirely
by the expected neutral coalescence noise around the expected coalescence time
$2B(b)N$. This can be be found analytically, plugging-in our predictions for
$B(b)$. This allows us to calculate $R_\text{coal}^2$ to ballpark the theoretic
variance that is capable of being explained, assuming this coalescence-only
noise process (see Supplementary Materials \ref{supp:r2-coal}). We note that
selection is expected to \emph{decrease} the variance in coalescence times
beyond a rescaled effective population size implies; our $R_\text{coal}^2$
would be an underestimate under selection.

@@

Early genome-wide implementations of the classic BGS model fit well, but lead
to unusually high estimates of the mutation rate \parencite{McVicker2009-ax}.
This lead to the hypothesis that these models could be absorbing the signal of
positive selection \parencite{Enard2014-kz}, though other work has found a
limited role for hitchhiking
\parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022}. While our simulation
results suggest estimates of the mutation rate from linked selection models are
biased, we still our estimates were in rough accordance with pedigree-based
estimates \parencite{Kong2013-fc,Tian2019-so}.

We find across all populations, our mutation rate estimates from CADD-based
models are roughly consistent with pedigree-based estimates (Figure
\ref{fig:figure-4}A), consistent with recent work \parencite{Murphy2022-sj}.
Our full-track CADD 6\% model estimates the mutation rate as $\widehat{\mu} =
1.56 \times 10^{-8}$ for YRI, $1.63 \times 10^{-8}$ for CEU, and $1.60 \times
10^{-8}$ for CHB reference samples (Supplementary Table \ref{supp:tbl-r2}). As
expected, the sparse-track CADD model mutation rate estimates are identical
between the B' and classic BGS methods (Figure \ref{fig:figure-4}A top row).

However, mutation rate estimates for feature-based annotation models are less
concordant with pedigree-based estimates. First, mutation rate estimates under
our implementation of classic BGS theory are an order of magnitude below the
expected range (Figure \ref{fig:figure-4}A top row). We observe similar
behavior when we use the classic BGS model to fit full-coverage annotation
models (Figure \ref{fig:figure-4}A bottom row). This behavior is consistent
with classic BGS theory being unable to fit the DFE to features under weak
constraint (e.g. introns, UTRs, and the ``other" feature), and most compensate 
by estimating an unusually low mutation rate.

Second, we noticed that across all populations and sparse/full tracks, the CADD
6\% model consistently led to slightly higher mutation rates than the CADD 8\%
model (Figure \ref{fig:figure-4}A bottom row; Supplementary Table
\ref{supp:tbl-r2}). This pattern was observed in Murphy et al. too (2023;
Appendix 1, Figure 16). This behavior suggests a non-identifiability issue
between a slightly higher per-basepair mutation rate and annotation tracks that
contain more conserved sequence. This is expected from theory, since both
classic BGS and SC16 models only depend on mutation rate through the compound
parameter $\mu L$, where $L$ is the length of the conserved segment. Even
though our method is much more robust to including non-conserved regions like
introns, we still observe this non-identifiability issue.

Finally, we note that mutation rate estimates in the Feature Priority model are
unreasonably high ($\widehat{\mu} \approx 3 \times 10^{-8}$), reminiscent of
the high mutation rate estimates found under McVicker et al.'s
(\citeyear{McVicker2009-ax}) model. While both our and Murphy et al.'s CADD and
PhastCons-based models alleviate this issue, it is worth discussing why this
could occur. We gain some insight from comparing the estimated mutation rates
between our Feature and PhastCons Priority annotation models, which both
contain the exact same number of basepairs, but these are assigned to different
feature classes based on the priority of overlapping features. That one of
these models is our best-fitting model, and the other our worst indicates the
sensitivity these models have when features classes have heterogeneous DFEs.
CADD-based models fit better in part due to their fine-scale resolution of
selective effects across the genome. While ideally we would fit a CADD model
with different features corresponding to the different percentiles of
pathogenicity, these features are on the basepair scale and thus too
memory-intensive for our method to currently accommodate.

@@

    \caption{(A) Observed and predicted diversity under the CADD 6\% full-track
        model. Once scaled by average diversity, populations differ little in
        their predicted diversity levels (colors). Additionally, we show
        summaries of CADD density and recombination rate along the chromosome
        below. (B) Predicted $B$ and observed $\pi$ for each window. The
        residual variance is in the vicinity of the levels expected from the
        $R_\text{LOCO}^2$ calculation. (C) Evidence that negative selection
    signal remains in CADD 6\% residuals (YRI shown). Residuals are plotted
against the average LoF selection coefficient across genes in a megabase window
(estimated by \cite{Agarwal2023-un}).}

@@

While we have previously shown that out-sample predicted diversity matches
observed diversity (Section \ref{sec:annotation}), we also investigated whether
there were regions of the genome that were particularly poor-fitting. Comparing
predicted against observed diversity along chromosomes (see Supplementary
Materials Section \ref{supp:chrom-fit}) we find a close correspondence
consistent with the high $R_\text{LOCO}^2$. However, we find a few large (tens
of megabases) regions with systematically poorer fit (Supplementary Materials
Figure \ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one
such region on the short arm of chromosome 2, from 30 Mbp to 60Mbp.
Interestingly, predicted diversity closely follows the peaks and troughs of
this region. We note that a small region within this stretch had been found by
a genome-wide scan for associative overdominance \parencite{Gilbert2020-aw}.

Another model diagnostic is to inspect whether observations are systematically
different from predictions. We confirm the finding of \textcite{Murphy2022-sj}
that regions predicted to experience little reduction in diversity due to
background selection (i.e. $B\approx 1$) have higher diversity than predicted
(Figure \ref{fig:figure-5}B, orange line). They suggested that this could
reflect ancient introgression between archaic humans and ancestors of
contemporary humans. Despite this aberration, the variance around observed and
predicted diversity levels falls very close to what we would expect under the
theoretic coalescence-noise-only expectation ($R_\text{coal}^2$).

Earlier, we found that the goodness-of-fit and DFE estimates of our
feature-based annotation models are sensitive to how basepairs are grouped into
feature classes. Our models are in essence fitting a single distribution across
sites within a feature class, yet there may be local heterogeneity among sites
that is poorly captured by this approach. We investigated this possibility by
looking for lingering selection signal in our model residuals. First, we
inspected whether there was a relationship between the fraction of CADD 2\% and
6\% basepairs and the residual across megabase windows (Supplementary Figure
\label{suppfig:resid-cadd}), finding a negative significant relationship in
both cases. Moreover, our model over-predicted diversity in windows containing
more CADD 2\% basepairs moreso than CADD 6\%, consistent with heterogeneity in
site pathogenicity being poorly fit by our model. However, the total residual
variance explained is $R^2 = 0.3\%$ and $R^2 = 0.9\%$ for the CADD 2\% and 6\%
tracks respectively, suggesting only modest amount of selection signal remains.

Since our method does not include the possible effects of linked positive
selection, we might expect windows containing hard or soft sweeps would have
systematically lower diversity levels than predicted. Using the locations of
soft and hard sweeps detected using a machine learning approach
\parencite{Schrider2017-yx,Kern2018-jd}, we tested whether the residuals of the
CADD 6\% model containing sweeps were systematically different than those not
containing sweeps. We find no significant difference between the magnitude of
residuals of windows containing sweeps versus those that do not (Supplementary
Materials Figure \ref{suppfig:resid-sweeps}; Kolmogorov--Smirnov p-value =
0.71).

@@

Since our B' method is based upon quantitative genetic models of linked
selection that explicitly incorporates the substitution process, fitting our
model to genomic diversity data leads to additional predictions. In particular,
our model predicts (1) the substitution rates per segment, (2) the fitness
variance per segment, and (3) the total number of strongly deleterious
mutations per diploid genome per generation. Given predicted substitution rates
for each feature can be compared to estimated sequence divergence between
species, this offers a qualitative check our model is producing sensible
results.

To compare our models' predicted substitution rates against observed
divergence, we estimated sequence divergence on the human lineage using a
multiple alignment of five primates for each feature in our feature-based
models (Methods XXX). Since our simulations show that mutation rate estimates
can be biased, we predicted substitution rates under a fixed mutation rate of
$\mu = 1.5 \times 10^{-8}$. Fixing the mutation rate also allows us to more
easily compare the predictions across our feature-based models. Unfortunately,
a careful comparison between our predictions and observed divergence rates is
hindered by considerable uncertainty in the rate of the molecular clock,
generation times, and the human-chimpanzee divergence time. We assume a
generation time of 29 years, and calculate the sequence divergence implied by
our predicted substitution rates over a range of divergence times, from 6 Mya
to 12 Mya \parencite{Moorjani2016-tb,Nachman2000-te,Yi2002-pw,Steiper2006-xx}.

Our predictions are qualitatively consistent with the observed divergence along
the human lineage (Figure \ref{fig:figure-7}). As expected, the predicted
substitution rate in features unlikely to be under selective constraint is very
close to the mutation rate. Throughout, we report our substitution rates as a
percent relative to the total mutation rate, $\mu$ (here fixed to $1.5 \times
10^{-8}$). In our PhastCons Priority model, introns and UTRs are estimated to
have have substitution rates that are $93.6\%$ of the mutation rate, and the
``other" feature is estimated at $97.8\%$.  When we use the constrained grid,
these change to $97.85\%$ and $99.98\%$ respectively. In our PhyloFit
estimates, the divergence rate in introns is $92.44\%$, 5' UTRs $75.5\%$, and
3' UTRs $72.65\%$. By contrast, our predicted substitution rates for PhastCons
regions are much lower: $0.48\%$ for the default grid, and $1.65\%$ for the
constrained grid. For comparison, the PhyloFit estimates for PhastCons regions
is much higher, at $51.23\%$. Since this would include beneficial substitutions
that would not be predicted by our model, this suggests a very high proportion
of beneficial substitutions XXX check
